### 2024-11-15 proposal_EDA/deGuzman_IDS702Projectproposal_template.qmd
update proposal templatre

---
title: "Project Proposal"
author: "Sizhe Chen, Naikyah Dhariwala, Peter de Guzman, Jennifer Li"
subtitle: "Due November 17 at 11:59pm"
format: pdf
---

#### Load Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
```

# Dataset 1 (top choice)

**Data source:** The 2020 Collaborative Multi-racial Post-Election Survey (CMPS)

**Brief description:** The survey has an unweighted sample size of 17,545 respondents. Each row represents a unique respondent. The data is collected through an online survey with 800 unique questions. The survey's main focus is on attitudes about the 2020 Election and candidates, experiences with racism, policy attitudes, immigration, and personal experiences with civic engagement across many facets of American life.

The 2020 CMPS was offered in English, Spanish, Chinese (simplified), Chinese (traditional), Korean, Vietnamese, Arabic, Urdu, Farsi and Haitian Creole. The survey dataset includes Black, White, Latino and Asian respondents as well as additional oversamples of respondents from hard-to-reach populations including, Afro-Latinos, Black immigrants, Native Americans, Native Hawaiians, Muslims and people who identify as LGBTQ.

**Research question 1:** Are Americans who believe their race is important to their identity more likely to be interested in politics than those that are not?

-   Outcome variable (include the name/description and type of variable): The outcome variable is an ordinal variable labeled Q29. The survey question is "Some people are very interested in politics while other people can't stand politics how about you? are you...". The response options are in a five-point scale from "Not at all interested in politics" to "Very interested in politics."

**Research question 2:**

-   Outcome variable (include the name/description and type of variable):

**Load the data and provide a `glimpse()`**:

```{r data-1, warning=FALSE}
rm(list=ls())
load("/Users/pdeguz01/Documents/git/Data/2020 CMPS/DS0001/39096-0001-Data.rda")
data <- da39096.0001
data %>% slice_head(n = 100) %>% glimpse()
```

**Exploratory Plots:**

```{r plots-1}
wtd_tableQ29 <- data %>% count(Q29, wt=WEIGHT)
wtd_tableQ29 <- wtd_tableQ29 %>% mutate('pct' = (n/sum(n, na.rm=TRUE)))

ggplot(wtd_tableQ29, aes(x=Q29, y = pct)) +
  geom_bar(stat="identity", fill = "orange") +
  geom_text(aes(label=percent(pct)), vjust=1.6, color="white", size =3.5) +
  labs(title = "Reported Interest in Politics", 
       x = "Level of Interest", y = "Percent (Weighted)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

contingency_table <- table(data$Q29, data$Q271)
contingency_df <- as.data.frame(contingency_table)
contingency_df <- contingency_df %>% mutate(Share = (Freq/sum(Freq)))

ggplot(contingency_df, aes(x = Var1, y = Var2, fill = Share)) +
  geom_tile() +
  labs(title="Heatmap of Interest in Politics and Racial Identity", x = "Interest in Politics", y = "Importance of Racial Identity") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_continuous(low = "white", high = "blue", labels = percent_format(scale = 100)) 
```

# Dataset 2

**Data source:**

**Brief description:**

**Research question 1:**

-   Outcome variable (include the name/description and type of variable):

**Research question 2:**

-   Outcome variable (include the name/description and type of variable):

**Load the data and provide a `glimpse()`**:

```{r data-2, warning=FALSE}

```

**Exploratory Plots:**

```{r plots-2}

```

#

# Dataset 3 (optional)

**Data source:**

**Brief description:**

**Research question 1:**

-   Outcome variable (include the name/description and type of variable):

**Research question 2:**

-   Outcome variable (include the name/description and type of variable):

**Load the data and provide a `glimpse()`**:

```{r data-3, warning=FALSE}

```

**Exploratory Plots:**

```{r plots-3}

```

#

# Team Charter

#### When will you meet as a team to work on the project components? Will these meetings be held in person or virtually?

#### What is your group policy on missing team meetings (e.g., how much advance notice should be provided)?

#### How will your team communicate (email, Slack, text messages)? What is your policy on appropriate response time (within a certain number of hours? Nights/weekends?)?

####