### 2024-12-02 Naki_BLR.qmd
First

---
title: "Naki_BLR"
format: html
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

The `echo: false` option disables the printing of code (only output is displayed).

# Loading dataset

```{r}
#| echo: false
# Set the file path
file_path <- "/Users/nakiyahdhariwala/StatsProject/39096-0001-Data.rda"

# Load the RDA file
load(file_path)

# List objects loaded into the environment
ls()

# Access the dataset (replace `da39096.0001` with the actual name found in `ls()`)
dataset <- da39096.0001  # Use the correct object name

```

The `echo: false` option disables the printing of code (only output is displayed).

```{r}
# Exploring the data

#colnames(dataset)
head(dataset)
```

# Variable selection

```{r}
required_cols <- c("Q1R5", # Outcome Variable; Stopping discrimination against racial/ethnic minorities (Binary: 0 = Not important, 1 = Important).
                   "S2_RACE_PRIME",   # Primary Race
                   "MAIN_ETH_QUOTA",  # Ethnic Quota
                   "S3",              # Orientation
                   "S3B",             # Gender
                   #"S4",              # Current state of residence
                   "S5_AGE",          # Age categories
                   "S13",             # Highest level of education
                   "S14",             # Community lived in

                   # Variables capturing experiences with racism
                   # Create stacked bar charts for this section to identify which variables we shoul look into further

                   "Q627",            # Treated unfairly or discriminated against

                   "Q629R1",          # Unfair treatment due to racial background/ethnicity
                   "Q629R2",          # Unfair treatment due to skin color
                   "Q629R3",          # Unfair treatment due to gender
                   "Q629R4",          # Unfair treatment due to sexuality/orientation
                   "Q629R5",          # Unfair treatment due to immigration status
                   "Q629R6",          # Unfair treatment due to religion
                   "Q629R7",          # Unfair treatment due to accent

                   "Q633",            # Discrimination impact on life satisfaction

                   # Religion-related variables
                   "Q58R1",            # Catholic
                   "Q58R2",            # Protestant
                   "Q58R3",            # Other Christian
                   "Q58R4",            # Jewish
                   "Q58R5",            # Muslim
                   "Q58R6",            # Hindu
                   "Q58R7",            # Buddhist
                   "Q58R8",            # Mormon/LDS
                   "Q58R9",            # Folk Religion
                   "Q58R10",           # Daoist
                   "Q58R11",           # Shamanism
                   "Q58R12",           # Ancestral Religion
                   "Q58R13",           # Spiritual but not religious
                   "Q58R14",           # Atheist
                   "Q58R15")            # Agnostic

# Create a new dataframe with selected variables
dataset_new <- dataset[, required_cols]
head(dataset_new)

```

# Convert variables

```{r}
# Convert the required columns to factors
dataset_new[required_cols] <- lapply(dataset_new[required_cols], as.factor)

# Verify that the columns have been successfully converted to factors
str(dataset_new[required_cols])
head(dataset_new)
```

```{r}
# Convert individual religion columns to numeric
for (i in 1:15) {
  col_name <- paste0("Q58R", i, "_numeric")
  dataset_new[[col_name]] <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", dataset_new[[paste0("Q58R", i)]]))
}

# Create a new Religion variable
dataset_new$Religion <- "None"
dataset_new$Religion[dataset_new$Q58R1_numeric == 1] <- "Catholic"
dataset_new$Religion[dataset_new$Q58R2_numeric == 1] <- "Protestant"
dataset_new$Religion[dataset_new$Q58R3_numeric == 1] <- "Other Christian"
dataset_new$Religion[dataset_new$Q58R4_numeric == 1] <- "Jewish"
dataset_new$Religion[dataset_new$Q58R5_numeric == 1] <- "Muslim"
dataset_new$Religion[dataset_new$Q58R6_numeric == 1] <- "Hindu"
dataset_new$Religion[dataset_new$Q58R7_numeric == 1] <- "Buddhist"
dataset_new$Religion[dataset_new$Q58R8_numeric == 1] <- "Mormon/LDS"
dataset_new$Religion[dataset_new$Q58R9_numeric == 1] <- "Folk Religion"
dataset_new$Religion[dataset_new$Q58R10_numeric == 1] <- "Daoist"
dataset_new$Religion[dataset_new$Q58R11_numeric == 1] <- "Shamanism"
dataset_new$Religion[dataset_new$Q58R12_numeric == 1] <- "Ancestral Religion"
dataset_new$Religion[dataset_new$Q58R13_numeric == 1] <- "Spiritual but not religious"
dataset_new$Religion[dataset_new$Q58R14_numeric == 1] <- "Atheist"
dataset_new$Religion[dataset_new$Q58R15_numeric == 1] <- "Agnostic"

# Check if an individual reported multiple religious affiliations 
dataset_new$multiple_religions <- rowSums(dataset_new[, grep("_numeric", names(dataset_new))] == 1)
check <- table(dataset_new$multiple_religions)

# Handle multiple religious affiliations
dataset_new$Religion[rowSums(dataset_new[, grep("_numeric", names(dataset_new))] == 1) > 1] <- "Multiple Religions"

# Convert Religion to a factor
dataset_new$Religion <- factor(dataset_new$Religion,
                               levels = c("Catholic", "Protestant", "Other Christian", "Jewish", "Muslim",
                                          "Hindu", "Buddhist", "Mormon/LDS", "Folk Religion",
                                          "Daoist", "Shamanism", "Ancestral Religion",
                                          "Spiritual but not religious", "Atheist", "Agnostic", 
                                          "Multiple Religions", "None"))

# Create a table of counts for the new Religion variable
table(dataset_new$Religion)

```

```{r}
# Remove redundant religion columns (Q58R1 to Q58R15 and their numeric versions)
redundant_cols <- c(
  "Q58R1", "Q58R2", "Q58R3", "Q58R4", "Q58R5", "Q58R6", "Q58R7", "Q58R8", 
  "Q58R9", "Q58R10", "Q58R11", "Q58R12", "Q58R13", "Q58R14", "Q58R15",
  "Q58R1_numeric", "Q58R2_numeric", "Q58R3_numeric", "Q58R4_numeric", 
  "Q58R5_numeric", "Q58R6_numeric", "Q58R7_numeric", "Q58R8_numeric", 
  "Q58R9_numeric", "Q58R10_numeric", "Q58R11_numeric", "Q58R12_numeric", 
  "Q58R13_numeric", "Q58R14_numeric", "Q58R15_numeric"
)

# Remove redundant columns from the dataset
dataset_clean <- dataset_new[, !(names(dataset_new) %in% redundant_cols)]

# Take a look at the updated dataset
head(dataset_clean)
```

# Missing values

```{r}
# Check the number of missing values for each variable
missing_summary <- colSums(is.na(dataset_clean))
missing_summary

# Finding proportion of Missing Data
missing_proportion <- colSums(is.na(dataset_clean)) / nrow(dataset_clean)
missing_proportion

# Finding total missing data proportion
overall_missing <- sum(is.na(dataset_clean)) / (nrow(dataset_clean) * ncol(dataset_clean))
overall_missing

# Omiting missing data (Not gonna do it for now)
# dataset_clean <- na.omit(dataset_clean)
# head(dataset_clean)
```

We can see that 9180 values are missing from every column that relates to questions of unfair treatment. Upon further investigation we find that missing values account for approximately 52% of the total column data. For now, we decide not to do anything and proceed further with the analysis

```{r}
head(dataset_new)
head(dataset_clean)
```

# Cleaned version

Make one chart showing the outcome distribution

```{r}
# (Outcome) Distribution of the proportion of people who view "Stopping discrimination against racial/ethnic minorities" as an important issue
table(dataset_clean$Q1R5) / nrow(dataset_clean)

table(dataset_clean$S2_RACE_PRIME)
table(dataset_clean$MAIN_ETH_QUOTA)

```
```{r}
# List of columns to be transformed
cols <- c("Q629R1", "Q629R2", "Q629R3", "Q629R4", "Q629R5", "Q629R6", "Q629R7")

# Convert factors to numeric (0 or 1) based on levels
dataset_clean$Q629R1_racialbackground <- ifelse(as.numeric(dataset_clean$Q629R1) == 1, 1, 0)
dataset_clean$Q629R2_skincolor <- ifelse(as.numeric(dataset_clean$Q629R2) == 1, 1, 0)
dataset_clean$Q629R3_gender <- ifelse(as.numeric(dataset_clean$Q629R3) == 1, 1, 0)
dataset_clean$Q629R4_orientation <- ifelse(as.numeric(dataset_clean$Q629R4) == 1, 1, 0)
dataset_clean$Q629R5_immigration <- ifelse(as.numeric(dataset_clean$Q629R5) == 1, 1, 0)
dataset_clean$Q629R6_religion <- ifelse(as.numeric(dataset_clean$Q629R6) == 1, 1, 0)
dataset_clean$Q629R7_accent <- ifelse(as.numeric(dataset_clean$Q629R7) == 1, 1, 0)

# Check the distribution of the newly created binary columns
for (col in c("Q629R1_racialbackground", "Q629R2_skincolor", "Q629R3_gender", 
              "Q629R4_orientation", "Q629R5_immigration", "Q629R6_religion", 
              "Q629R7_accent")) {
  print(col)  # Print the column name
  print(table(dataset_clean[[col]]))  # Access the column dynamically
}
```

```{r}
# Checking for levels in the model
level <- sapply(dataset_clean, function(x) length(unique(x)))
level

dataset_clean$interaction_term <- interaction(dataset_clean$S2_RACE_PRIME, dataset_clean$MAIN_ETH_QUOTA)

model <- glm(Q1R5 ~ S2_RACE_PRIME +   # Primary Race
                    MAIN_ETH_QUOTA +  # Ethnic Quota
                   S3 +              # Orientation
                   S3B +             # Gender
                   S5_AGE +          # Age categories
                   S13 +             # Highest level of education
                   S14 +             # Community lived in

                   # Variables capturing experiences with racism
                  Q627 +           # Treated unfairly or discriminated against
                  Religion +
               interaction_term,
                  #Q629R1,          # Unfair treatment due to racial background/ethnicity
                  #Q629R2 +          # Unfair treatment due to skin color
                  #Q629R3 +          # Unfair treatment due to gender
                  #Q629R4 +          # Unfair treatment due to sexuality/orientation
                  #Q629R5 +          # Unfair treatment due to immigration status
                  #Q629R6 +          # Unfair treatment due to religion
                  #Q629R7 +          # Unfair treatment due to accent
                  # 
                  #Q633,             # Discrimination impact on life satisfaction
                   data = dataset_clean, family = binomial)            
summary(model)

#prop.table(table(dataset_clean$Q629R1))
#prop.table(table(dataset_clean$Q629R2))
# Repeat for other variables
```

```{r}
step_model <- step(model, direction = "both")
step_model
```