### 2024-12-02 Naki_EDA_Exploration.qmd

---
title: "EDA_Naki"
format: html
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

# Loading dataset

```{r}
#| echo: false
# Set the file path
file_path <- "/Users/nakiyahdhariwala/StatsProject/39096-0001-Data.rda"

# Load the RDA file
load(file_path)

# List objects loaded into the environment
ls()

# Access the dataset (replace `da39096.0001` with the actual name found in `ls()`)
dataset <- da39096.0001  # Use the correct object name

```

The `echo: false` option disables the printing of code (only output is displayed).

```{r}
# Exploring the data

#colnames(dataset)
head(dataset)
```

# Variable selection

```{r}
required_cols <- c("Q1R5", # Outcome Variable; Stopping discrimination against racial/ethnic minorities (Binary: 0 = Not important, 1 = Important).
                   "S2_RACE_PRIME",   # Primary Race
                   # "MAIN_ETH_QUOTA",  # Ethnic Quota
                   "S3",              # Orientation
                   "S3B",             # Gender
                   "S4",              # Current state of residence
                   "S5_AGE",          # Age categories
                   "S13",             # Highest level of education
                   "S14",             # Community lived in

                   # Variables capturing experiences with racism
                   # Create stacked bar charts for this section to identify which variables we shoul look into further 
                   "Q627",            # Treated unfairly or discriminated against
                   # "Q628",            # Location of discrimination
                   "Q629R1",          # Unfair treatment due to racial background/ethnicity
                   "Q629R2",          # Unfair treatment due to skin color
                   "Q629R3",          # Unfair treatment due to gender
                   "Q629R4",          # Unfair treatment due to sexuality/orientation
                   "Q629R5",          # Unfair treatment due to immigration status
                   "Q629R6",          # Unfair treatment due to religion
                   "Q629R7",          # Unfair treatment due to accent

                   "Q633",            # Discrimination impact on life satisfaction
                   # Religion-related variables
                   "Q58R1",            # Catholic
                   "Q58R2",            # Protestant
                   "Q58R3",            # Other Christian
                   "Q58R4",            # Jewish
                   "Q58R5",            # Muslim
                   "Q58R6",            # Hindu
                   "Q58R7",            # Buddhist
                   "Q58R8",            # Mormon/LDS
                   "Q58R9",            # Folk Religion
                   "Q58R10",           # Daoist
                   "Q58R11",           # Shamanism
                   "Q58R12",           # Ancestral Religion
                   "Q58R13",           # Spiritual but not religious
                   "Q58R14",           # Atheist
                   "Q58R15")            # Agnostic

# Create a new dataframe with selected variables
dataset_new <- dataset[, required_cols]
head(dataset_new)

```

# Convert variables

```{r}
# Convert the required columns to factors
dataset_new[required_cols] <- lapply(dataset_new[required_cols], as.factor)

# Verify that the columns have been successfully converted to factors
str(dataset_new[required_cols])
head(dataset_new)
```

```{r}
# Convert individual religion columns to numeric
for (i in 1:15) {
  col_name <- paste0("Q58R", i, "_numeric")
  dataset_new[[col_name]] <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", dataset_new[[paste0("Q58R", i)]]))
}

# Create a new Religion variable
dataset_new$Religion <- "None"
dataset_new$Religion[dataset_new$Q58R1_numeric == 1] <- "Catholic"
dataset_new$Religion[dataset_new$Q58R2_numeric == 1] <- "Protestant"
dataset_new$Religion[dataset_new$Q58R3_numeric == 1] <- "Other Christian"
dataset_new$Religion[dataset_new$Q58R4_numeric == 1] <- "Jewish"
dataset_new$Religion[dataset_new$Q58R5_numeric == 1] <- "Muslim"
dataset_new$Religion[dataset_new$Q58R6_numeric == 1] <- "Hindu"
dataset_new$Religion[dataset_new$Q58R7_numeric == 1] <- "Buddhist"
dataset_new$Religion[dataset_new$Q58R8_numeric == 1] <- "Mormon/LDS"
dataset_new$Religion[dataset_new$Q58R9_numeric == 1] <- "Folk Religion"
dataset_new$Religion[dataset_new$Q58R10_numeric == 1] <- "Daoist"
dataset_new$Religion[dataset_new$Q58R11_numeric == 1] <- "Shamanism"
dataset_new$Religion[dataset_new$Q58R12_numeric == 1] <- "Ancestral Religion"
dataset_new$Religion[dataset_new$Q58R13_numeric == 1] <- "Spiritual but not religious"
dataset_new$Religion[dataset_new$Q58R14_numeric == 1] <- "Atheist"
dataset_new$Religion[dataset_new$Q58R15_numeric == 1] <- "Agnostic"

# Convert Religion to a factor
dataset_new$Religion <- factor(dataset_new$Religion,
                               levels = c("Catholic", "Protestant", "Other Christian", "Jewish", "Muslim",
                                          "Hindu", "Buddhist", "Mormon/LDS", "Folk Religion",
                                          "Daoist", "Shamanism", "Ancestral Religion",
                                          "Spiritual but not religious", "Atheist", "Agnostic", "None"))

# Create a table of counts for the new Religion variable
table(dataset_new$Religion)

```

```{r}
# Remove redundant religion columns (Q58R1 to Q58R15 and their numeric versions)
redundant_cols <- c(
  "Q58R1", "Q58R2", "Q58R3", "Q58R4", "Q58R5", "Q58R6", "Q58R7", "Q58R8", 
  "Q58R9", "Q58R10", "Q58R11", "Q58R12", "Q58R13", "Q58R14", "Q58R15",
  "Q58R1_numeric", "Q58R2_numeric", "Q58R3_numeric", "Q58R4_numeric", 
  "Q58R5_numeric", "Q58R6_numeric", "Q58R7_numeric", "Q58R8_numeric", 
  "Q58R9_numeric", "Q58R10_numeric", "Q58R11_numeric", "Q58R12_numeric", 
  "Q58R13_numeric", "Q58R14_numeric", "Q58R15_numeric"
)

# Remove redundant columns from the dataset
dataset_clean <- dataset_new[, !(names(dataset_new) %in% redundant_cols)]

# Take a look at the updated dataset
head(dataset_clean)
```

# Missing values

```{r}
# Check the number of missing values for each variable
missing_summary <- colSums(is.na(dataset_clean))
missing_summary

# Finding proportion of Missing Data
missing_proportion <- colSums(is.na(dataset_clean)) / nrow(dataset_clean)
missing_proportion

# Finding total missing data proportion
overall_missing <- sum(is.na(dataset_new)) / (nrow(dataset_new) * ncol(dataset_new))
overall_missing

# Omiting missing data (Not gonna do it for now)
# dataset_clean <- na.omit(dataset_clean)
# head(dataset_clean)
```

We can see that 9180 values are missing from every column that relates to questions of unfair treatment. Upon further investigation we find that missing values account for approximately 52% of the total column data. For now, we decide not to do anything and proceed further with the analysis

# Cleaned version

```{r}
# (Outcome) Distribution of the proportion of people who view "Stopping discrimination against racial/ethnic minorities" as an important issue
table(dataset_new$Q1R5) / nrow(dataset_new)
```

```{r}
#Summarize counts and proportions for every variable

#lapply(dataset_new[, -1], table)

```

```{r}
# Function to create a summary table with counts and proportions
create_summary_table <- function(variable, var_name) {
  counts <- table(variable)
  if (length(counts) == 0) {
    return(data.frame(Variable = character(), Level = character(), Count = numeric(), Proportion = numeric()))
  }
  proportions <- prop.table(counts)
  summary_df <- data.frame(
    Variable = var_name,
    Level = names(counts),
    Count = as.numeric(counts),
    Proportion = round(as.numeric(proportions), 4)
  )
  return(summary_df)
}

# Apply the function to each variable of interest in the cleaned dataset
categorical_vars <- names(dataset_clean)

# Initialize an empty data frame to store all summary tables
full_summary_table <- data.frame(Variable = character(), Level = character(), Count = numeric(), Proportion = numeric())

# Iterate through each variable and generate the summary table
for (var in categorical_vars) {
  summary_table <- create_summary_table(dataset_clean[[var]], var)
  full_summary_table <- rbind(full_summary_table, summary_table)
}

# Print the full summary table
full_summary_table

```

| **Variable** | **Level** | **Count** | **Proportion** |
|------------------|--------------------|------------------|------------------|
| Q1R5 | \(0\) NO TO: Stopping discrimination against racial/ethnic minorities | 5728 | 0.6848 |
| Q1R5 | \(1\) Stopping discrimination against racial/ethnic minorities | 2637 | 0.3152 |
| S2_RACE_PRIME | \(1\) White | 1310 | 0.1566 |
| S2_RACE_PRIME | \(2\) Hispanic or Latino | 1836 | 0.2195 |
| S2_RACE_PRIME | \(3\) Black or African American | 2733 | 0.3267 |
| S2_RACE_PRIME | \(4\) Asian American | 1834 | 0.2192 |
| S2_RACE_PRIME | \(5\) American Indian/Native American | 357 | 0.0427 |
| S2_RACE_PRIME | \(6\) Arab, Middle Eastern or North African | 145 | 0.0173 |
| S2_RACE_PRIME | \(7\) Native Hawaiian | 39 | 0.0047 |
| S2_RACE_PRIME | \(8\) Not Hawaiian, but other Pacific Islander | 111 | 0.0133 |
| MAIN_ETH_QUOTA | \(1\) White | 955 | 0.1142 |
| MAIN_ETH_QUOTA | \(2\) Latino | 2007 | 0.2399 |
| MAIN_ETH_QUOTA | \(3\) Black | 2578 | 0.3082 |
| MAIN_ETH_QUOTA | \(4\) AAPI | 1898 | 0.2269 |
| MAIN_ETH_QUOTA | \(5\) Native American | 552 | 0.066 |
| MAIN_ETH_QUOTA | \(6\) MENA | 176 | 0.021 |
| MAIN_ETH_QUOTA | \(7\) Native Hawaiian | 199 | 0.0238 |
| S3 | \(01\) Straight, that is, not gay or lesbian | 7037 | 0.8412 |
| S3 | \(02\) Gay or lesbian | 391 | 0.0467 |
| S3 | \(03\) Bisexual | 587 | 0.0702 |
| S3 | \(04\) Something else (Specify) | 207 | 0.0247 |
| S3 | \(88\) Refused | 90 | 0.0108 |
| S3 | \(99\) I don | 53 | 0.0063 |
| S3B | \(1\) Man | 3388 | 0.405 |
| S3B | \(2\) Woman | 4844 | 0.5791 |
| S3B | \(3\) Non-binary | 109 | 0.013 |
| S3B | \(4\) Something else (Specify) | 24 | 0.0029 |
| S4 | \(01\) Alaska | 18 | 0.0022 |
| S4 | \(02\) Alabama | 120 | 0.0143 |
| S4 | \(03\) Arizona | 207 | 0.0247 |
| S4 | \(04\) Arkansas | 74 | 0.0088 |
| S4 | \(05\) California | 1456 | 0.1741 |
| S4 | \(06\) Colorado | 136 | 0.0163 |
| S4 | \(07\) Connecticut | 77 | 0.0092 |
| S4 | \(08\) Delaware | 28 | 0.0033 |
| S4 | \(09\) District of Columbia | 43 | 0.0051 |
| S4 | \(10\) Florida | 677 | 0.0809 |
| S4 | \(11\) Georgia | 349 | 0.0417 |
| S4 | \(12\) Hawaii | 84 | 0.01 |
| S4 | \(13\) Idaho | 15 | 0.0018 |
| S4 | \(14\) Illinois | 333 | 0.0398 |
| S4 | \(15\) Indiana | 99 | 0.0118 |
| S4 | \(16\) Iowa | 27 | 0.0032 |
| S4 | \(17\) Kansas | 39 | 0.0047 |
| S4 | \(18\) Kentucky | 74 | 0.0088 |
| S4 | \(19\) Louisiana | 112 | 0.0134 |
| S4 | \(20\) Maine | 8 | 0.001 |
| S4 | \(21\) Maryland | 208 | 0.0249 |
| S4 | \(22\) Massachusetts | 142 | 0.017 |
| S4 | \(23\) Michigan | 182 | 0.0218 |
| S4 | \(24\) Minnesota | 79 | 0.0094 |
| S4 | \(25\) Mississippi | 68 | 0.0081 |
| S4 | \(26\) Missouri | 89 | 0.0106 |
| S4 | \(27\) Montana | 8 | 0.001 |
| S4 | \(28\) Nebraska | 24 | 0.0029 |
| S4 | \(29\) Nevada | 92 | 0.011 |
| S4 | \(30\) New Hampshire | 13 | 0.0016 |
| S4 | \(31\) New Jersey | 248 | 0.0296 |
| S4 | \(32\) New Mexico | 56 | 0.0067 |
| S4 | \(33\) New York | 593 | 0.0709 |
| S4 | \(34\) North Carolina | 258 | 0.0308 |
| S4 | \(35\) North Dakota | 11 | 0.0013 |
| S4 | \(36\) Ohio | 235 | 0.0281 |
| S4 | \(37\) Oklahoma | 88 | 0.0105 |
| S4 | \(38\) Oregon | 74 | 0.0088 |
| S4 | \(39\) Pennsylvania | 231 | 0.0276 |
| S4 | \(40\) Rhode Island | 16 | 0.0019 |
| S4 | \(41\) South Carolina | 99 | 0.0118 |
| S4 | \(42\) South Dakota | 13 | 0.0016 |
| S4 | \(43\) Tennessee | 135 | 0.0161 |
| S4 | \(44\) Texas | 862 | 0.103 |
| S4 | \(45\) Utah | 46 | 0.0055 |
| S4 | \(46\) Vermont | 7 | 0.0008 |
| S4 | \(47\) Virginia | 210 | 0.0251 |
| S4 | \(48\) Washington | 202 | 0.0241 |
| S4 | \(49\) West Virginia | 19 | 0.0023 |
| S4 | \(50\) Wisconsin | 78 | 0.0093 |
| S4 | \(51\) Wyoming | 3 | 0.0004 |
| S5_AGE | \(1\) 16-17 | 0 | 0.0 |
| S5_AGE | \(2\) 18-29 | 2150 | 0.257 |
| S5_AGE | \(3\) 30-39 | 1639 | 0.1959 |
| S5_AGE | \(4\) 40-49 | 1376 | 0.1645 |
| S5_AGE | \(5\) 50-59 | 1289 | 0.1541 |
| S5_AGE | \(6\) 60-69 | 1289 | 0.1541 |
| S5_AGE | \(7\) 70+ | 622 | 0.0744 |
| S13 | \(1\) Grades 1-8 | 58 | 0.0069 |
| S13 | \(2\) Some High School, but did not graduate | 292 | 0.0349 |
| S13 | \(3\) High School graduate or GED | 1335 | 0.1596 |
| S13 | \(4\) Some college | 1834 | 0.2192 |
| S13 | \(5\) Associates, 2-year degree | 960 | 0.1148 |
| S13 | \(6\) Bachelors, 4-year degree | 2210 | 0.2642 |
| S13 | \(7\) Post-graduate degree | 1676 | 0.2004 |
| S14 | \(1\) Large urban area | 2517 | 0.3009 |
| S14 | \(2\) Large suburb near large city | 2793 | 0.3339 |
| S14 | \(3\) Small suburb near small town or city | 1331 | 0.1591 |
| S14 | \(4\) Small town or small city | 1130 | 0.1351 |
| S14 | \(5\) Rural area | 594 | 0.071 |
| Q627 | \(1\) Yes | 8365 | 1.0 |
| Q627 | \(2\) No | 0 | 0.0 |
| Q629R1 | \(0\) NO TO: Racial background or ethnicity | 2947 | 0.3523 |
| Q629R1 | \(1\) Racial background or ethnicity | 5418 | 0.6477 |
| Q629R2 | \(0\) NO TO: Skin color | 4040 | 0.483 |
| Q629R2 | \(1\) Skin color | 4325 | 0.517 |
| Q629R3 | \(0\) NO TO: Gender | 5340 | 0.6384 |
| Q629R3 | \(1\) Gender | 3025 | 0.3616 |
| Q629R4 | \(0\) NO TO: Sexuality or sexual orientation | 7219 | 0.863 |
| Q629R4 | \(1\) Sexuality or sexual orientation | 1146 | 0.137 |
| Q629R5 | \(0\) NO TO: Immigration status | 7718 | 0.9227 |
| Q629R5 | \(1\) Immigration status | 647 | 0.0773 |
| Q629R6 | \(0\) NO TO: Religion | 7299 | 0.8726 |
| Q629R6 | \(1\) Religion | 1066 | 0.1274 |
| Q629R7 | \(0\) NO TO: Accent, regardless of whether or not you have an accent | 6881 | 0.8226 |
| Q629R7 | \(1\) Accent, regardless of whether or not you have an accent | 1484 | 0.1774 |
| Q633 | \(1\) A lot | 997 | 0.1192 |
| Q633 | \(2\) Some | 2002 | 0.2393 |
| Q633 | \(3\) A little | 2964 | 0.3543 |
| Q633 | \(4\) Not at all | 1908 | 0.2281 |
| Q633 | \(5\) Don | 426 | 0.0509 |
| Q633 | \(6\) Refused | 68 | 0.0081 |
| Religion | Catholic | 1484 | 0.1774 |
| Religion | Protestant | 1901 | 0.2273 |
| Religion | Other Christian | 931 | 0.1113 |
| Religion | Jewish | 144 | 0.0172 |
| Religion | Muslim | 265 | 0.0317 |
| Religion | Hindu | 190 | 0.0227 |
| Religion | Buddhist | 231 | 0.0276 |
| Religion | Mormon/LDS | 50 | 0.0060 |
| Religion | Folk Religion | 15 | 0.0018 |
| Religion | Daoist | 11 | 0.0013 |
| Religion | Shamanism | 17 | 0.0020 |
| Religion | Ancestral Religion | 51 | 0.0061 |
| Religion | Spiritual but not religious | 787 | 0.0941 |
| Religion | Atheist | 444 | 0.0531 |
| Religion | Agnostic | 601 | 0.0718 |
| Religion | None | 1243 | 0.1486 |

# Visualization

```{r}
library(ggplot2)
library(dplyr)

```

# `Q627` vs `Q1R5`: Experience of Racism vs Stopping Discrimination

```{r}
ggplot(dataset_clean, aes(x = Q627, fill = Q1R5)) +
  geom_bar(position = "fill") +
  labs(
    title = "Q627 vs Q1R5",
    y = "Proportion",
    x = "Q627 (Experience of Racism)",
    fill = "Q1R5 (Stopping Discrimination)"
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal()

```

# `MAIN_ETH_QUOTA` vs `Q1R5`: Ethnic Quota vs Stopping Discrimination

```{r}
ggplot(dataset_clean, aes(x = MAIN_ETH_QUOTA, fill = Q1R5)) +
  geom_bar(position = "fill") +
  labs(
    title = "MAIN_ETH_QUOTA vs Q1R5",
    y = "Proportion",
    x = "MAIN_ETH_QUOTA (Ethnic Quota)",
    fill = "Q1R5 (Stopping Discrimination)"
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    plot.title = element_text(hjust = 0.5)
  )
```

# `S3B` vs `Q1R5`: Gender vs Stopping Discrimination

```{r}
ggplot(dataset_clean, aes(x = S3B, fill = Q1R5)) +
  geom_bar(position = "fill") +
  labs(
    title = "S3B vs Q1R5",
    y = "Proportion",
    x = "S3B (Gender)",
    fill = "Q1R5 (Stopping Discrimination)"
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    plot.title = element_text(hjust = 0.5)
  )
```

# `S2_RACE_PRIME` vs `Q1R5`: Primary Race vs Stopping Discrimination

```{r}
ggplot(dataset_clean, aes(x = S2_RACE_PRIME, fill = Q1R5)) +
  geom_bar(position = "fill") +
  labs(
    title = "S2_RACE_PRIME vs Q1R5",
    y = "Proportion",
    x = "S2_RACE_PRIME (Primary Race)",
    fill = "Q1R5 (Stopping Discrimination)"
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal()+
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    plot.title = element_text(hjust = 0.5)
  )

```

# `S5_AGE` vs `Q1R5`: Age vs Stopping Discrimination

```{r}
ggplot(dataset_clean, aes(x = S5_AGE, fill = Q1R5)) +
  geom_bar(position = "fill") +
  labs(
    title = "S5_AGE vs Q1R5",
    y = "Proportion",
    x = "S5_AGE (Age Group)",
    fill = "Q1R5 (Stopping Discrimination)"
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal()+
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    plot.title = element_text(hjust = 0.5)
  )

```

# `S14` vs `Q1R5`: Community Type vs Stopping Discrimination

```{r}
ggplot(dataset_clean, aes(x = S14, fill = Q1R5)) +
  geom_bar(position = "fill") +
  labs(
    title = "S14 vs Q1R5",
    y = "Proportion",
    x = "S14 (Community Type)",
    fill = "Q1R5 (Stopping Discrimination)"
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal()+
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    plot.title = element_text(hjust = 0.5)
  )

```

# Plot for `S13` (Education Level)

```{r}
ggplot(dataset_clean, aes(x = S13, fill = Q1R5)) +
  geom_bar(position = "fill") +
  labs(
    title = "S13 vs Q1R5",
    y = "Proportion",
    x = "S13 (Highest Level of Education)",
    fill = "Q1R5 (Stopping Discrimination)"
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    plot.title = element_text(hjust = 0.5)
  )

```

# Plot for `Q58` (Religion Variables)

```{r}
ggplot(dataset_clean, aes(x = Religion, fill = Q1R5)) +
  geom_bar(position = "fill") +
  labs(
    title = "Q58 vs Q1R5",
    y = "Proportion",
    x = "Religion",
    fill = "Q1R5 (Stopping Discrimination)"
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
    plot.title = element_text(hjust = 0.5)
  )

```

# Plot for `Q633` (Impact of Discrimination on Life Satisfaction)

```{r}
ggplot(dataset_clean, aes(x = Q633, fill = Q1R5)) +
  geom_bar(position = "fill") +
  labs(
    title = "Q633 vs Q1R5",
    y = "Proportion",
    x = "Q633 (Discrimination Impact on Life Satisfaction)",
    fill = "Q1R5 (Stopping Discrimination)"
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    plot.title = element_text(hjust = 0.5)
  )
```

# Plot for `Q629` (Unfair Treatment Variables)

```{r}
# Create a sample dataset to use for plotting
plot_data <- data.frame(
  Variable = c(rep("Q629R1", 2), rep("Q629R2", 2), rep("Q629R3", 2), rep("Q629R4", 2), rep("Q629R5", 2), rep("Q629R6", 2), rep("Q629R7", 2)),
  Level = c("(0) NO TO: Racial background or ethnicity", "(1) Racial background or ethnicity",
            "(0) NO TO: Skin color", "(1) Skin color",
            "(0) NO TO: Gender", "(1) Gender",
            "(0) NO TO: Sexuality or sexual orientation", "(1) Sexuality or sexual orientation",
            "(0) NO TO: Immigration status", "(1) Immigration status",
            "(0) NO TO: Religion", "(1) Religion",
            "(0) NO TO: Accent, regardless of whether or not you have an accent", "(1) Accent, regardless of whether or not you have an accent"),
  Count = c(2947, 5418, 4040, 4325, 5340, 3025, 7219, 1146, 7718, 647, 7299, 1066, 6881, 1484),
  Proportion = c(0.3523, 0.6477, 0.483, 0.517, 0.6384, 0.3616, 0.863, 0.137, 0.9227, 0.0773, 0.8726, 0.1274, 0.8226, 0.1774)
)

# Plot for each variable
ggplot(plot_data, aes(x = Variable, y = Proportion, fill = Level)) +
  geom_bar(stat = "identity", position = "fill") +
  labs(title = "Proportions of Different Variables by Level",
       x = "Variable",
       y = "Proportion") +
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# chi-square

```{r}
dataset_clean$S5_AGE <- droplevels(dataset_clean$S5_AGE)

# Create a summary table for all categorical variables
variables <- c("Q1R5", "S2_RACE_PRIME", "MAIN_ETH_QUOTA", "S3", "S3B", "S5_AGE", "S13", "Q627", "Q629R1", "Q629R2", "Q629R3", "Q629R4", "Q629R5", "Q629R6", "Q629R7", "Q633", "Religion")

# Chi-square tests for each variable
chi_square_results <- lapply(variables, function(var) {
  chisq.test(table(dataset_clean[[var]], dataset_clean$Q1R5))
})

# Print chi-square results
names(chi_square_results) <- variables
print(chi_square_results)

```

| Variable | Chi-Square Statistic | Degrees of Freedom (df) | P-Value | Test Type |
|---------------|---------------|---------------|---------------|---------------|

|               |        |     |            |                            |
|---------------|--------|-----|------------|----------------------------|
| S2_RACE_PRIME | 416.06 | 7   | \< 2.2e-16 | Pearson's Chi-squared test |

|                |        |     |            |                            |
|----------------|--------|-----|------------|----------------------------|
| MAIN_ETH_QUOTA | 383.38 | 6   | \< 2.2e-16 | Pearson's Chi-squared test |

|     |        |     |         |                            |
|-----|--------|-----|---------|----------------------------|
| S3  | 11.184 | 5   | 0.04786 | Pearson's Chi-squared test |

|     |        |     |           |                            |
|-----|--------|-----|-----------|----------------------------|
| S3B | 31.259 | 3   | 7.497e-07 | Pearson's Chi-squared test |

|        |        |     |        |                            |
|--------|--------|-----|--------|----------------------------|
| S5_AGE | 6.2621 | 5   | 0.2815 | Pearson's Chi-squared test |

|     |        |     |         |                            |
|-----|--------|-----|---------|----------------------------|
| S13 | 13.215 | 6   | 0.03975 | Pearson's Chi-squared test |

|      |     |     |     |                            |
|------|-----|-----|-----|----------------------------|
| Q627 | NaN | 1   | NA  | Pearson's Chi-squared test |

|  |  |  |  |  |
|---------------|---------------|---------------|---------------|---------------|
| Q629R1 | 114.83 | 1 | \< 2.2e-16 | Pearson's Chi-squared test with Yates' correction |

|  |  |  |  |  |
|---------------|---------------|---------------|---------------|---------------|
| Q629R2 | 72.719 | 1 | \< 2.2e-16 | Pearson's Chi-squared test with Yates' correction |

|        |         |     |        |                                                   |
|---------------|---------------|---------------|---------------|---------------|
| Q629R3 | 0.41214 | 1   | 0.5209 | Pearson's Chi-squared test with Yates' correction |

|        |        |     |        |                                                   |
|---------------|---------------|---------------|---------------|---------------|
| Q629R4 | 1.3169 | 1   | 0.2511 | Pearson's Chi-squared test with Yates' correction |

|        |       |     |           |                                                   |
|---------------|---------------|---------------|---------------|---------------|
| Q629R5 | 11.48 | 1   | 0.0007036 | Pearson's Chi-squared test with Yates' correction |

|  |  |  |  |  |
|---------------|---------------|---------------|---------------|---------------|
| Q629R6 | 20.113 | 1 | 7.301e-06 | Pearson's Chi-squared test with Yates' correction |

|        |        |     |        |                                                   |
|---------------|---------------|---------------|---------------|---------------|
| Q629R7 | 2.6287 | 1   | 0.1049 | Pearson's Chi-squared test with Yates' correction |

|      |        |     |           |                            |
|------|--------|-----|-----------|----------------------------|
| Q633 | 70.581 | 5   | 7.759e-14 | Pearson's Chi-squared test |

|          |        |     |           |                            |
|----------|--------|-----|-----------|----------------------------|
| Religion | 84.974 | 15  | 8.507e-12 | Pearson's Chi-squared test |

```{r}
# Model Fitting - Logistic Regression
# Logistic regression to predict stopping discrimination (Q1R5)

# Create interaction term between MAIN_ETH_QUOTA and Religion
dataset_clean$interaction_term <- interaction(dataset_clean$MAIN_ETH_QUOTA, dataset_clean$Religion)

```

```{r}
dataset_clean <- droplevels(dataset_clean)
dataset_clean <- dataset_clean %>%
  mutate(across(where(is.character), as.factor))
# Refitting the logistic regression model without Q627
model <- glm(Q1R5 ~ S2_RACE_PRIME + MAIN_ETH_QUOTA + S3 + S3B + S5_AGE + S13 + Q629R1 + Q633 + Religion + interaction_term,
             data = dataset_clean,
             family = binomial())

# Summary of the model
summary(model)
```

### **Interaction Terms**

1.  **Experiences with Racism × Perceptions of Discrimination**:

    -   Interaction between **Q627** (personal experience) and **Q619-Q626** (perceived discrimination toward groups).

2.  **Race/Ethnicity × Social Prioritization**:

    -   Interaction between **S2_RACE** and policy prioritization for racial equity issues (e.g., Q117-Q125).

3.  **Racism Experience × Leadership Perceptions**:

    -   Interaction between **Q627** and perceived efficacy of racial/ethnic leadership (Q159-Q161).

4.  **Interaction (Q627 × S2_RACE)**:

    -   Investigates how the impact of discrimination varies by racial/ethnic group, directly addressing intersectionality.

### **Key Recommendations**

1.  Focus on **Q627**, **Q629R1**, and **S2_RACE** as the core predictors.

2.  Use **interaction terms** (e.g., Q627 × S2_RACE) to explore differences across racial/ethnic groups.

3.  Include **Q633** to capture the subjective impact of discrimination.

4.  Use **perceptions of systemic discrimination (Q619-Q626)** to contextualize personal experiences.

```{r}
```

```{r echo: false}
# Creating an interaction term
# Create interaction term
dataset$Interaction <- interaction(dataset$Q627, dataset$S2_RACE)

                   # "S2_RACE_PRIME",   # Primary Race
                   # "MAIN_ETH_QUOTA",  # Ethnic Quota
                   # "S3",              # How you think of yourself
                   # "S3B",             # Gender
                   # "S4",              # Current state of residence
                   # # "S5",            # Year of birth (commented out)
                   # "S5_AGE",          # Age categories
                   # "S13",             # Highest level of education
                   # "S14",             # Community lived in

                   # # Variables capturing experiences with racism
                   # "Q627",            # Treated unfairly or discriminated against
                   # # "Q628",            # Location of discrimination
                   # "Q629R1",          # Unfair treatment due to racial background/ethnicity
                   # "Q629R2",          # Unfair treatment due to skin color
                   # "Q629R3",          # Unfair treatment due to gender
                   # "Q629R4",          # Unfair treatment due to sexuality/orientation
                   # "Q629R5",          # Unfair treatment due to immigration status
                   # "Q629R6",          # Unfair treatment due to religion
                   # "Q629R7",          # Unfair treatment due to accent
                   # "Q633",            # Discrimination impact on life satisfaction
                   # "Q619_Q626R2",     # Perceptions of discrimination against Blacks
                   # "Q619_Q626R3",     # Perceptions of discrimination against Asian Americans
                   # "Q619_Q626R4",     # Perceptions of discrimination against Native Americans
                   # "Q619_Q626R5",     # Perceptions of discrimination against Immigrants
                   # "Q619_Q626R6")    # Perceptions of discrimination against Latinos

# Logistic regression model
model <- glm(
  Q1R5 ~ S2_RACE_PRIME + MAIN_ETH_QUOTA + S3 + S3B + S4 + S5_AGE + S13 + S14 + 
   Q627 + Q629R1 + Q629R2 + Q629R3 + Q629R4 + Q629R5 + Q629R6 + Q629R7 +
    Q633 +
     Q619_Q626R2 + Q619_Q626R3 + Q619_Q626R4 + Q619_Q626R5 + Q619_Q626R6,
  family = binomial(link = "logit"),
  data = dataset_new
)

# Summarize model
summary(model)
```

```{r}
# Check levels of all variables used in the model
sapply(dataset[, c("S2_RACE_PRIME", "MAIN_ETH_QUOTA", "S3", "S3B", "S4", "S5_AGE", "S13", "S14", 
                   "Q627", "Q629R1", "Q629R2", "Q629R3", "Q629R4", "Q629R5", "Q629R6", "Q629R7",
                   "Q633", "Q619_Q626R2", "Q619_Q626R3", "Q619_Q626R4", "Q619_Q626R5", "Q619_Q626R6")], 
       function(x) if(is.factor(x)) levels(x) else unique(x))

# Identify single-level factors
single_level_vars <- names(Filter(function(x) length(unique(x)) <= 1, dataset_new))
print(single_level_vars)

```

```{r}
{r}
# Example: Stacked bar plot for Q627
library(ggplot2)
ggplot(dataset_new, aes(x = Q627, fill = Q1R5)) +
  geom_bar(position = "fill") +
  labs(title = "Q627 vs Q1R5", y = "Proportion", x = "Q627 (Experience of Racism)")

ggplot(dataset_new, aes(x = MAIN_ETH_QUOTA, fill = Q1R5)) +
  geom_bar(position = "fill") +
  labs(title = "Q627 vs Q1R5", y = "Proportion", x = "Q627 (Experience of Racism)")
```