### 2024-12-02 peter_EDA.qmd
adding files

---
title: "peter_EDA"
format: html
author: "Peter de Guzman"
editor: visual
---

Conducting EDA on the 2020 CMPS dataset for Research Question 1

# Loading data

# Selecting Variables

# Cleaning Variables

# Analysis of Missing Data

# Save cleaned data

# Creating summary table

# Running chi-square tests

# Fitting OLR Model

# Fitting MLR Model