### 2024-12-02 peter_plots.qmd

---
title: "peter_plots"
format: html
editor: visual
---

##

# Loading data

# plot 1:

# plot 2:

# plot 3: