### 2024-12-02 peter_EDA.qmd
updating EDA qmd

---
title: "peter_EDA"
format: html
author: "Peter de Guzman"
editor: visual
---

Conducting EDA on the 2020 CMPS dataset for Research Question 1

# Loading data
```{r, echo=FALSE}
file_path <- "/Users/pdeguz01/Documents/git/Data/2020 CMPS/DS0001/39096-0001-Data.rda"
# Load data 
load(file_path)
data <- da39096.0001
rm(da39096.0001)
```

# Selecting Variables

Need to include the outcome variable, which is an ordinal variable Q29.
Main predictor variable is Q271 "How important is being [RACIAL GROUP] to your identity?"
```{r}
required_cols <- c("Q29", # Outcome Variable; interest in politics
                   "Q271", # Predictor Variable; importance of race to identity
                   "S3", # Orientation
                   "S3B", # Gender
                   "S5_AGE", # Age Category
                   "S13", # Educational Attainment
                   "S14", # Community Type
                   "S7", # Nativity
                   "Q21", # Party ID
                   "S2_M", # Multiracial

                   # Religion-related variables
                   "Q58R1",            # Catholic
                   "Q58R2",            # Protestant
                   "Q58R3",            # Other Christian
                   "Q58R4",            # Jewish
                   "Q58R5",            # Muslim
                   "Q58R6",            # Hindu
                   "Q58R7",            # Buddhist
                   "Q58R8",            # Mormon/LDS
                   "Q58R9",            # Folk Religion
                   "Q58R10",           # Daoist
                   "Q58R11",           # Shamanism
                   "Q58R12",           # Ancestral Religion
                   "Q58R13",           # Spiritual but not religious
                   "Q58R14",           # Atheist
                   "Q58R15",            # Agnostic

                   # Variables capturing experiences with racism
                   "Q627", # Treated unfairly or discriminated against ,
                   "Q628", # Location of discrimination
                   "Q629R1" # Unfair treatment due to racial background/ethnicity
                   )

# Create new df
data_filtered <- data[, required_cols]
head(data_filtered)

```

# Cleaning Variables

```{r}
# Convert the required columns to factors
data_filtered[required_cols] <- lapply(data_filtered[required_cols], as.factor)

# Verify that the columns have been successfully converted to factors
str(data_filtered[required_cols])
head(data_filtered)
```
Cleaning and Recoding Religion Variables
```{r}
# Convert individual religion columns to numeric
for (i in 1:15) {
  col_name <- paste0("Q58R", i, "_numeric")
  dataset_new[[col_name]] <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", dataset_new[[paste0("Q58R", i)]]))
}

# Create a new Religion variable
dataset_new$Religion <- "None"
dataset_new$Religion[dataset_new$Q58R1_numeric == 1] <- "Catholic"
dataset_new$Religion[dataset_new$Q58R2_numeric == 1] <- "Protestant"
dataset_new$Religion[dataset_new$Q58R3_numeric == 1] <- "Other Christian"
dataset_new$Religion[dataset_new$Q58R4_numeric == 1] <- "Jewish"
dataset_new$Religion[dataset_new$Q58R5_numeric == 1] <- "Muslim"
dataset_new$Religion[dataset_new$Q58R6_numeric == 1] <- "Hindu"
dataset_new$Religion[dataset_new$Q58R7_numeric == 1] <- "Buddhist"
dataset_new$Religion[dataset_new$Q58R8_numeric == 1] <- "Mormon/LDS"
dataset_new$Religion[dataset_new$Q58R9_numeric == 1] <- "Folk Religion"
dataset_new$Religion[dataset_new$Q58R10_numeric == 1] <- "Daoist"
dataset_new$Religion[dataset_new$Q58R11_numeric == 1] <- "Shamanism"
dataset_new$Religion[dataset_new$Q58R12_numeric == 1] <- "Ancestral Religion"
dataset_new$Religion[dataset_new$Q58R13_numeric == 1] <- "Spiritual but not religious"
dataset_new$Religion[dataset_new$Q58R14_numeric == 1] <- "Atheist"
dataset_new$Religion[dataset_new$Q58R15_numeric == 1] <- "Agnostic"

# Convert Religion to a factor
dataset_new$Religion <- factor(dataset_new$Religion,
                               levels = c("Catholic", "Protestant", "Other Christian", "Jewish", "Muslim",
                                          "Hindu", "Buddhist", "Mormon/LDS", "Folk Religion",
                                          "Daoist", "Shamanism", "Ancestral Religion",
                                          "Spiritual but not religious", "Atheist", "Agnostic", "None"))

# Create a table of counts for the new Religion variable
table(dataset_new$Religion)

```

Removing depracated Religion variables
```{r}
# Remove redundant religion columns (Q58R1 to Q58R15 and their numeric versions)
redundant_cols <- c(
  "Q58R1", "Q58R2", "Q58R3", "Q58R4", "Q58R5", "Q58R6", "Q58R7", "Q58R8", 
  "Q58R9", "Q58R10", "Q58R11", "Q58R12", "Q58R13", "Q58R14", "Q58R15",
  "Q58R1_numeric", "Q58R2_numeric", "Q58R3_numeric", "Q58R4_numeric", 
  "Q58R5_numeric", "Q58R6_numeric", "Q58R7_numeric", "Q58R8_numeric", 
  "Q58R9_numeric", "Q58R10_numeric", "Q58R11_numeric", "Q58R12_numeric", 
  "Q58R13_numeric", "Q58R14_numeric", "Q58R15_numeric"
)

# Remove redundant columns from the dataset
dataset_clean <- dataset_new[, !(names(dataset_new) %in% redundant_cols)]

# Take a look at the updated dataset
head(dataset_clean)
```

Cleaning and Recoding Other Variables

# Analysis of Missing Data
```{r}

```

# Save cleaned data
```{r}

```

# Creating summary table
```{r}

```