### 2024-12-03 peter_plots.qmd
updating recoding script

---
title: "peter_plots"
format: html
editor: visual
---

## 
# Packages
```{r}
library(tidyverse)
library(janitor)
library(ggplot2)
library(scales)
```

# Loading data
```{r}
file_path = "/Users/pdeguz01/Documents/git/Data/2020 CMPS/Output/2020cmps_clean.RData"
load(file_path)
```

# plot 1:
Race and importance of being a member of racial group to identity (Q271)
```{r}

data_summary <- dataset_clean %>%
  group_by(S2_RACE_PRIME, Q271) %>%
  summarise(count = n()) %>%
  mutate(proportion = count / sum(count) * 100)  # Calculate proportion for each race

races <- unique(data_summary$S2_RACE_PRIME)

# Loop through each race and generate a plot
for (race_value in races) {
  plot_data <- data_summary %>% filter(S2_RACE_PRIME == race_value)

  p <- ggplot(plot_data, aes(x = Q271, y = proportion, fill = Q271)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(title = paste("Proportions for", race_value, "Respondents"), x = "Importance of Race to Identity", y = "Proportion", fill = "Rating") + 
    scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentages
    theme_minimal() +
    theme(
      axis.text = element_text(size = 12),  # Customize axis text size
      plot.title = element_text(size = 14, hjust = 0.5)  # Center title
    )

  # Save each plot as a separate image file (optional)
  #ggsave(paste0("plot_", race_value, ".png"), plot = p, width = 6, height = 4)

  # Print the plot (optional, if you want to view them in the R console)
  print(p)
}

```

# plot 2:
Race and interest in politics (Q29)
```{r}

```

# plot 3:
White/non-white and importance of race to identity (Q271)

# plot:
White/non-white and interest in politics (Q29)

# plot 3:
College Experience and importance of being a member of racial group to identity (Q271)
```{r}
data_summary <- dataset_clean %>%
  group_by(educ, Q271) %>%
  summarise(count = n()) %>%
  mutate(proportion = count / sum(count) * 100)  # Calculate proportion for each race

plot_somecollege <- data_summary %>% filter(educ == "Some college experience") %>% ggplot(aes(x = Q271, y = proportion, fill = Q271)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(title = paste("Proportions for Respondents with Some college Experience"), x = "Importance of Race to Identity", y = "Proportion", fill = "Rating") + 
    scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentages
    theme_minimal() +
    theme(
      axis.text = element_text(size = 12),  # Customize axis text size
      plot.title = element_text(size = 14, hjust = 0.5)  # Center title
    )

print(plot_somecollege)

plot_nocollege <- data_summary %>% filter(educ == "No college experience") %>% ggplot(aes(x = Q271, y = proportion, fill = Q271)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(title = paste("Proportions for Respondents with No college Experience"), x = "Importance of Race to Identity", y = "Proportion", fill = "Rating") + 
    scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentages
    theme_minimal() +
    theme(
      axis.text = element_text(size = 12),  # Customize axis text size
      plot.title = element_text(size = 14, hjust = 0.5)  # Center title
    )

print(plot_nocollege)

```

# plot 4:
College Experience and interest in politics (Q29)
```{r}

```

# plot 4:
Age and importance of being a member of racial group to identity (Q271)
```{r}

```

# plot 5:
Age and interest in politics (Q29)
```{r}

```

# plot 6:
Rurality and importance of being a member of racial group to identity (Q271)
```{r}

```

# plot 7:
Rurality and interest in politics
```{r}

```