### 2024-12-14 RQ2_final.qmd
Look at RQ2

---
title: "RQ2_final"
format: html
editor: visual

execute: 
  warning: false
---

### Loading Libraries

```{r echo = FALSE}
# Load necessary libraries
library(dplyr)   # Provides %>% and mutate
library(tidyverse)  # Optionally load the entire tidyverse if needed
library(pROC)
library(vcd)
library(car)
library(knitr)
library(gt)
library(caret)
```

```{r echo = FALSE, results = FALSE}
# Read the dataset
file_path <- "39096-0001-Data.rda"

load(file_path)          # Load the RDA file
ls()                     # List objects loaded into the environment
dataset <- da39096.0001  # Access the dataset

head(dataset)
ncol(dataset)
nrow(dataset)
```

## Pre-processing Data

```{r echo = FALSE, results=FALSE}
# Define the required columns
required_cols <- c(
  # Outcome Variable: Stopping discrimination against racial/ethnic minorities
  "Q1R5",

  # Demographic Variables
  "S2_RACE_PRIME",   # Primary Race
  "S3",              # SexualOrientation
  "S3B",             # Gender
  "S5_AGE",          # Age categories
  "S13",             # Highest level of education
  "S14",             # Community lived in

  # Variables capturing experiences with racism
  "Q629R1",          # Unfair treatment due to racial background/ethnicity
  "Q629R2",          # Unfair treatment due to skin color
  "Q629R3",          # Unfair treatment due to gender
  "Q629R4",          # Unfair treatment due to sexuality/orientation
  "Q629R5",          # Unfair treatment due to immigration status
  "Q629R6",          # Unfair treatment due to religion
  "Q629R7",          # Unfair treatment due to accent

  "Q627",            # Treated unfairly or discriminated against because of your race, ethnicity, gender, sexuality, being an immigrant, religious heritage or having an accent?
  "Q633",            # Discrimination impact on life satisfaction

  # Perceptions of discrimination
  "Q619_Q626R1",     # Perceptions of discrimination against Whites
  "Q619_Q626R2",     # Perceptions of discrimination against Blacks
  "Q619_Q626R3",     # Perceptions of discrimination against Asian Americans
  "Q619_Q626R4",     # Perceptions of discrimination against Native Americans
  "Q619_Q626R5",     # Perceptions of discrimination against Immigrants
  "Q619_Q626R6",     # Perceptions of discrimination against Latinos
  "Q619_Q626R7",     # Perceptions of discrimination against GaysAndLesbians
  "Q619_Q626R8"      # Perceptions of discrimination against Muslims
)
```

```{r echo=FALSE, results = FALSE}
# Create a new dataframe with selected variables
dataset_new <- dataset[, required_cols]

# Recode the outcome variable 'Q1R5' for readability
dataset_new$Q1R5 <- factor(
  ifelse(
    dataset_new$Q1R5 == "(0) NO TO: Stopping discrimination against racial/ethnic minorities", 
    "Not important",
    "Important"  # Remaining cases are "Important"
  ),
  levels = c("Not important", "Important") # Define factor levels
)

# Define the columns for perceptions of discrimination
perception_cols <- c(
  "Q619_Q626R1",  # Perceptions of discrimination against Whites
  "Q619_Q626R2",  # Perceptions of discrimination against Blacks
  "Q619_Q626R3",  # Perceptions of discrimination against Asian Americans
  "Q619_Q626R4",  # Perceptions of discrimination against Native Americans
  "Q619_Q626R5",  # Perceptions of discrimination against Immigrants
  "Q619_Q626R6",  # Perceptions of discrimination against Latinos
  "Q619_Q626R7",  # Perceptions of discrimination against GaysAndLesbians
  "Q619_Q626R8",  # Perceptions of discrimination against Muslims
  "Q633"          # Discrimination impact on life satisfaction
)

# Recode 'Don' as 'Don't know' and ensure columns remain factors
dataset_new[perception_cols] <- lapply(dataset_new[perception_cols], function(x) {
  x <- as.character(x)  # Temporarily convert to character for recoding
  x <- ifelse(x == "(5) Don", "(5) Don't know", x)  # Recode 'Don' to 'Don't know'
  factor(x)  # Convert back to factor
})

# Define required columns to convert to factors (if not already defined)
# Assuming `required_cols` is previously defined; otherwise, specify explicitly
dataset_new[required_cols] <- lapply(dataset_new[required_cols], as.factor)

# Verify the structure of the dataset
str(dataset_new[required_cols])

# Display the first few rows to check the changes
head(dataset_new)
```

```{r echo=FALSE, results=FALSE}
# Data Cleaning and Recoding for dataset_new
dataset_clean <- dataset_new |> 
  mutate(
    # Education recoding
    educ = case_when(
      S13 == "(1) Grades 1-8" ~ "No college experience",
      S13 == "(2) Some High School, but did not graduate" ~ "No college experience",
      S13 == "(3) High School graduate or GED" ~ "No college experience",
      S13 == "(4) Some college" ~ "Some college experience",
      S13 == "(5) Associates, 2-year degree" ~ "College graduate",
      S13 == "(6) Bachelors, 4-year degree" ~ "College graduate",
      S13 == "(7) Post-graduate degree" ~ "College graduate"
    ),
    # Sexual orientation recoding
    sex_orientation = case_when(
      S3 == "(01) Straight, that is, not gay or lesbian" ~ "Straight",
      S3 == "(02) Gay or lesbian" ~ "Queer",
      S3 == "(03) Bisexual" ~ "Queer",
      S3 == "(04) Something else (Specify)" ~ "Queer",
      S3 == "(88) Refused" ~ "Refused/Don't Know",
      S3 == "(99) I don" ~ "Refused/Don't Know"
    ),
    # Urban vs rural recoding
    rural = case_when(
      S14 == "(1) Large urban area" ~ "Urban/suburban",
      S14 == "(2) Large suburb near large city" ~ "Urban/suburban",
      S14 == "(3) Small suburb near small town or city" ~ "Urban/suburban",
      S14 == "(4) Small town or small city" ~ "Urban/suburban",
      S14 == "(5) Rural area" ~ "Rural"
    ),
    # Race recoding
    race_recoded = case_when(
      S2_RACE_PRIME == "(1) White" ~ "White",
      S2_RACE_PRIME == "(2) Hispanic or Latino" ~ "Latino",
      S2_RACE_PRIME == "(3) Black or African American" ~ "Black",
      S2_RACE_PRIME == "(4) Asian American" ~ "Asian American",
      S2_RACE_PRIME == "(5) American Indian/Native American" ~ "AIAN",
      S2_RACE_PRIME == "(6) Arab, Middle Eastern or North African" ~ "Arab/MENA",
      S2_RACE_PRIME == "(7) Native Hawaiian" ~ "NHPI",
      S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "NHPI"
    ),
    # Age recoding
    age_recoded = case_when(
      S5_AGE == "(1) 16-17" ~ "16-17",
      S5_AGE == "(2) 18-29" ~ "18-29",
      S5_AGE == "(3) 30-39" ~ "30-39",
      S5_AGE == "(4) 40-49" ~ "40-49",
      S5_AGE == "(5) 50-59" ~ "50-59",
      S5_AGE == "(6) 60-69" ~ "60-69",
      S5_AGE == "(7) 70 +" ~ "70+"
    )
  ) |>
  mutate(
    # Converting variables to factors
    race_recoded = factor(race_recoded, levels = c("White", "Black", "Asian American", "Latino", "AIAN", "NHPI", "Arab/MENA")),
    age_recoded = factor(age_recoded, levels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+")),
    educ = factor(educ, levels = c("No college experience", "Some college experience", "College graduate")),
    rural = factor(rural, levels = c("Rural", "Urban/suburban")),
    sex_orientation = factor(sex_orientation, levels = c("Straight", "Queer", "Refused/Don't Know"))
  )

# Take a look at the updated dataset
head(dataset_clean)
```

```{r echo=FALSE, results=FALSE}
# Remove all redundant columns
redundant_columns <- c("Q629R1", "Q629R2", "Q629R3", "Q629R4", 
                       "Q629R5", "Q629R6", "Q629R7",
                       "S2_RACE_PRIME", "S13", "S3", "S5_AGE","S14")

dataset_clean <- dataset_clean[, !(names(dataset_clean) %in% redundant_columns)]

```

```{r echo=FALSE, results=FALSE}
# Rename remaining columns in dataset_clean
colnames(dataset_clean) <- c(
  "ImportanceOfStoppingDiscriminationAgainstMinorities",  # Q1R5
  "Gender",                                               # S3B
  "TreatedUnfairlyOrDiscriminatedAgainst",                # Q627
  "DiscriminationImpactOnLifeSatisfaction",              # Q633
  "PerceptionsOfDiscriminationAgainstWhites",            # Q619_Q626R1
  "PerceptionsOfDiscriminationAgainstBlacks",            # Q619_Q626R2
  "PerceptionsOfDiscriminationAgainstAsianAmericans",    # Q619_Q626R3
  "PerceptionsOfDiscriminationAgainstNativeAmericans",   # Q619_Q626R4
  "PerceptionsOfDiscriminationAgainstImmigrants",        # Q619_Q626R5
  "PerceptionsOfDiscriminationAgainstLatinos",           # Q619_Q626R6
  "PerceptionsOfDiscriminationAgainstGaysAndLesbians",   # Q619_Q626R7
  "PerceptionsOfDiscriminationAgainstMuslims",           # Q619_Q626R8
  "Education",                                           # educ
  "SexOrientation",                                      # sex_orientation
  "Rural",                                               # rural
  "Race",                                                # race_recoded
  "Age"                                                  # age_recoded
)

```

# Model Building

## Model 1: Main

```{r echo=FALSE, results=FALSE}
# Regression model
base_model <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ Gender +
               TreatedUnfairlyOrDiscriminatedAgainst + 
               DiscriminationImpactOnLifeSatisfaction +
               PerceptionsOfDiscriminationAgainstWhites + 
               PerceptionsOfDiscriminationAgainstBlacks +
               PerceptionsOfDiscriminationAgainstAsianAmericans + 
               PerceptionsOfDiscriminationAgainstNativeAmericans + 
               PerceptionsOfDiscriminationAgainstImmigrants + 
               PerceptionsOfDiscriminationAgainstLatinos + 
               PerceptionsOfDiscriminationAgainstGaysAndLesbians + 
               PerceptionsOfDiscriminationAgainstMuslims +
               Education + 
                 SexOrientation +
                 Rural +
                 Race +
                 Age,

             data = dataset_clean, family = "binomial")
```

## Model 2: Model with interaction term

```{r echo=FALSE, results=FALSE}
interaction_model <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ 
               TreatedUnfairlyOrDiscriminatedAgainst + 
               DiscriminationImpactOnLifeSatisfaction +
               PerceptionsOfDiscriminationAgainstWhites + 
               PerceptionsOfDiscriminationAgainstBlacks +
               PerceptionsOfDiscriminationAgainstAsianAmericans + 
               PerceptionsOfDiscriminationAgainstNativeAmericans + 
               PerceptionsOfDiscriminationAgainstImmigrants + 
               PerceptionsOfDiscriminationAgainstLatinos + 
               PerceptionsOfDiscriminationAgainstGaysAndLesbians + 
               PerceptionsOfDiscriminationAgainstMuslims +
               Education + 
                 SexOrientation +
                 Rural +
                 Age + 
                 Gender * Race,

             data = dataset_clean, family = "binomial")
```

#VIF Values

```{r}
# Check multicollinearity using VIF
vif_base <- vif(base_model)
vif_interaction <- vif(interaction_model)
```

|   | Normalized VIF: Base Model | Normalized VIF: Interaction Model |
|------------------------|------------------------|------------------------|
| Gender | 1.022 | 1.978 |
| Treated Unfairly or Discriminated Against | 1.125 | 1.126 |
| Discrimination Impact on Life Satisfaction | 1.036 | 1.037 |
| Perceptions Of Discrimination Against Whites | 1.105 | 1.106 |
| Perceptions Of Discrimination Against Blacks | 1.329 | 1.329 |
| Perceptions Of Discrimination Against Asian Americans | 1.308 | 1.308 |
| Perceptions Of Discrimination Against Native Americans | 1.224 | 1.224 |
| Perceptions Of Discrimination Against Immigrants | 1.357 | 1.359 |
| Perceptions Of Discrimination Against Latinos | 1.370 | 1.371 |
| Perceptions Of Discrimination Against Gays and Lesbians | 1.262 | 1.262 |
| Perceptions Of Discrimination Against Muslims | 1.341 | 1.342 |
| Education | 1.058 | 1.060 |
| Orientation | 1.042 | 1.044 |
| Community Type | 1.017 | 1.018 |
| Age | 1.028 | 1.030 |
| Race | 1.027 | 1.725 |
| Gender \* Race |  | 1.308 |

: Table 1: VIF Values

# Selecting the better model

## Calculating model's goodness-of-fit through deviance and chi-square test

```{r}
anova(base_model, interaction_model, test="Chisq")
```

## Calculating Odds Ratio, Confidence Interval, and p-values (base model)

```{r}
# Extract odds ratios
odds_ratios <- exp(coef(base_model))

# Get confidence intervals for coefficients
conf_intervals <- exp(confint(base_model))

# Extract p-values
p_values <- summary(base_model)$coefficients[, 4]

# Combine everything into a single data frame
results_base_model <- data.frame(
  OddsRatio = odds_ratios,
  LowerCI = conf_intervals[, 1],
  UpperCI = conf_intervals[, 2],
  PValue = p_values
)

# Display results
results_base_model
```

```{r}
# Verify the structure of the dataset
str(dataset_clean)
```

## Creating the analysis table

```{r}
names <- c("Intercept",
           "Gender: Women",
           "Gender: Non-Binary",
           "Gender: Something Else",
           "Treated Unfairly Or Discriminated Against: No", # Yes
           "Discrimination Impact On Life Satisfaction: Some", # A lot 
           "Discrimination Impact On Life Satisfaction: A Little",
           "Discrimination Impact On Life Satisfaction: Not At All",
           "Discrimination Impact On Life Satisfaction: Don't Know",
           "Discrimination Impact On Life Satisfaction: Refused",
           "Perceptions Of Discrimination Against Whites: Some", # A lot 
           "Perceptions Of Discrimination Against Whites: A Little",
           "Perceptions Of Discrimination Against Whites: None At All",
           "Perceptions Of Discrimination Against Whites: Don't Know",
           "Perceptions Of Discrimination Against Blacks: Some", # A lot 
           "Perceptions Of Discrimination Against Blacks: A Little",
           "Perceptions Of Discrimination Against Blacks: None At All",
           "Perceptions Of Discrimination Against Blacks: Don't Know",
           "Perceptions Of Discrimination Against Asian Americans: Some", # A lot 
           "Perceptions Of Discrimination Against Asian Americans: A Little",
           "Perceptions Of Discrimination Against Asian Americans: None At All",
           "Perceptions Of Discrimination Against Asian Americans: Don't Know",
           "Perceptions Of Discrimination Against Native Americans: Some", # A lot 
           "Perceptions Of Discrimination Against Native Americans: A Little",
           "Perceptions Of Discrimination Against Native Americans: None At All",
           "Perceptions Of Discrimination Against Native Americans: Don't Know",
           "Perceptions Of Discrimination Against Immigrants: Some", # A lot 
           "Perceptions Of Discrimination Against Immigrants: A Little",
           "Perceptions Of Discrimination Against Immigrants: None At All",
           "Perceptions Of Discrimination Against Immigrants: Don't Know",
           "Perceptions Of Discrimination Against Latinos: Some", # A lot 
           "Perceptions Of Discrimination Against Latinos: A Little",
           "Perceptions Of Discrimination Against Latinos: None At All",
           "Perceptions Of Discrimination Against Latinos: Don't Know",
           "Perceptions Of Discrimination Against Gays And Lesbians: Some", # A lot 
           "Perceptions Of Discrimination Against Gays And Lesbians: A Little",
           "Perceptions Of Discrimination Against Gays And Lesbians: None At All",
           "Perceptions Of Discrimination Against Gays And Lesbians: Don't Know",
           "Perceptions Of Discrimination Against Muslims: Some",# A lot 
           "Perceptions Of Discrimination Against Muslims: A Little",
           "Perceptions Of Discrimination Against Muslims: None At All",
           "Perceptions Of Discrimination Against Muslims: Don't Know",
           "Education: Some College Experience", # No College Experience
           "Education: College Graduate",
           "Orientation: Queer", #Straight
           "Orientation: Refused",
           "Community Type: Urban/Suburban", # Rural
           "Race: Black", # White
           "Race: Asian American",
           "Race: Latino",
           "Race: AIAN",
           "Race: NHPI",
           "Race: Arab/MENA",
           "Age: 30-39 Years", # 18-29
           "Age: 40-49 Years",
           "Age: 50-59 Years",
           "Age: 60-69 Years",
           "Age: 70+ Years"
           )
```

```         
```

```{r}
# Add row names as a new column
analysis_table <- cbind(RowName = names, results_base_model)

# Render the table
table <- analysis_table |>
  gt() |>
  tab_header(
    title = "Odds Ratios with Confidence Intervals and P-Values for base model"
  ) |>
  cols_label(
    RowName = "Description",  # Change the column header for the row names
    OddsRatio = "Odds Ratio",
    LowerCI = "Lower Confidence Interval",
    UpperCI = "Upper Confidence Interval",
    PValue = "P-Value"
  ) |>
  fmt_number(
    columns = c(OddsRatio, LowerCI, UpperCI),
    decimals = 4
  ) |>
  fmt_scientific(
    columns = c(PValue),
    decimals = 2  # Show p-values in scientific notation
  )

# View the table
table
```

```{r}
# Update the column names in the dataframe
colnames(analysis_table) <- c("Description", "Odds Ratio", "Lower Confidence Interval", 
                              "Upper Confidence Interval", "P-Value")

# View analysis table
analysis_table

# View the updated dataframe
head(analysis_table)
```

## Figure 1: Odds values

```{r}
# Filter for significant variables (p-value < 0.05)
significant_data <- analysis_table[analysis_table$`P-Value` < 0.05, ]

# Add a column for variable names to use in the plot (from the "Description" column)
significant_data <- significant_data |>
  mutate(Variable = Description)

# Create the plot
ggplot(significant_data, aes(x = `Odds Ratio`, y = Variable)) +
  geom_point(size = 4, color = "blue") +  # Add points for odds ratios
  geom_errorbarh(aes(xmin = `Lower Confidence Interval`, 
                     xmax = `Upper Confidence Interval`), 
                 height = 0.3, color = "black") +  # Add horizontal error bars
  geom_vline(xintercept = 1, linetype = "dashed", color = "red", size = 0.7) +  # Add a vertical reference line at Odds Ratio = 1
  labs(
    title = "Odds Ratio of Significant Variables",
    x = "Odds Ratio",
    y = "Variable"
  ) +
  theme_minimal() +  # Use a minimal theme for better aesthetics
  theme(
    axis.text.y = element_text(size = 6, color = "black"),  # Improve y-axis text readability
    axis.title.x = element_text(size = 12, face = "bold"),  # Highlight x-axis title
    axis.title.y = element_text(size = 12, face = "bold"),  # Highlight y-axis title
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),  # Center the title
    plot.margin = margin(1, 1, 1, 1, "cm"),  # Adjust margins
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),  # Reduce x-axis text size
    legend.text = element_text(size = 8),  # Reduce legend text size
    legend.title = element_text(size = 9)  # Reduce legend title size
  )

# Save the plot with larger dimensions
ggsave("Significant_Variables_Plot_Enhanced.png", width = 40, height = 20)

```

# Model Diagnostics and Evaluation

```{r}
# Generating a prediction as binary values (0 or 1), where:
## 1 indicates a predicted probability greater than 0.5 (positive class, "Important").
## 0 indicates a predicted probability less than or equal to 0.5 (negative class, "Not important").
model_prediction <- ifelse(predict(base_model, 
                                   newdata = dataset_clean, 
                                   type = "response") > 0.5, 1, 0)

# Converting the predictions factors
## 0 Not important, 1 Important
model_prediction_factor <- factor(model_prediction, 
                                  levels = c(0, 1), 
                                  labels = c("Not important", "Important"))  # Adjusting labels

# Convert the response variable to a factor with matching levels
response_factor <- factor(dataset_clean$ImportanceOfStoppingDiscriminationAgainstMinorities,
                          levels = c("Not important", "Important"))  # Ensuring matching order

# Checking if factors are identical (you can uncomment this to check the levels)
# levels(model_prediction_factor)
# levels(response_factor)

# Generate confusion matrix (if needed)
confusionMatrix(table(model_prediction_factor, response_factor),
                positive = "Important",
                mode = "everything")

```

|                                      |                   |               |
|--------------------------------------|-------------------|---------------|
| [**Predicted / Actual**]{.underline} | **Not Important** | **Important** |
| **Not Important**                    | 12361             | 3812          |
| **Important**                        | 666               | 706           |

: Table CONFUSION: Confusion Matrix for Base model

|                           | Metric  |
|---------------------------|---------|
| Sensitivity (Recall)      | 0.1563  |
| Specificity               | 0.9489  |
| Precision                 | 0.5146  |
| F1                        | 0.2397  |
| Balanced Accuracy         | 0.5526  |
| No Information Rate (NIR) | 0.7425  |
| Kappa                     | 0.1361  |
| Mcnemar's Test P-Value    | \<2e-16 |

: Table MODEL: Base Model Statistics

## ROC Curve (fig2)

```{r}
# Generate predicted probabilities (not just binary predictions) from the model
model_probabilities <- predict(base_model, newdata = dataset_clean, type = "response")

# Calculate ROC curve
roc_curve <- roc(response_factor, model_probabilities)

# Calculate AUC (Area Under the Curve)
auc_value <- auc(roc_curve)
print(paste("AUC: ", auc_value))

# Plot ROC curve
plot(roc_curve, main = "ROC Curve", col = "blue", lwd = 2)
legend("bottomright", legend = paste("AUC =", round(auc_value, 3)), col = "blue", lwd = 2)
```