### 2025-01-12 Project Proposal/_IDS702Projectproposal_Main.qmd

---
title: "Project Proposal"
author: "Peter de Guzman, Sizhe Chen, Jennifer Li, Nakiyah Dhariwala"
subtitle: "Due November 17 at 11:59pm"
format: pdf
editor: 
  markdown: 
    wrap: 72
---

#### Load Packages

```{r load-packages, message = FALSE, warning = FALSE, echo=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(readr)
library(ggridges)
library(readxl)
library(formattable)
library(scales)
```

# Dataset 1 (top choice)

**Data source:** The 2020 Collaborative Multi-racial Post-Election
Survey (CMPS)

**Brief description:** The survey has an unweighted sample size of
17,545 respondents. Each row represents a unique respondent. The data is
collected through an online survey with 800 unique questions. The
survey's main focus is on attitudes about the 2020 Election and
candidates, experiences with racism, policy attitudes, immigration, and
personal experiences with civic engagement across many facets of
American life.

The 2020 CMPS was offered in English, Spanish, Chinese (simplified),
Chinese (traditional), Korean, Vietnamese, Arabic, Urdu, Farsi and
Haitian Creole. The survey dataset includes Black, White, Latino and
Asian respondents as well as additional oversamples of respondents from
hard-to-reach populations including, Afro-Latinos, Black immigrants,
Native Americans, Native Hawaiians, Muslims and people who identify as
LGBTQ.

**Research question 1:** Are Americans who believe their race is
important to their identity more likely to be interested in politics
than those that are not?

-   Outcome variable (include the name/description and type of
    variable): The outcome variable is an ordinal variable labeled Q29.
    The survey question is "Some people are very interested in politics
    while other people can't stand politics how about you? are you...".
    The response options are in a five-point scale from "Not at all
    interested in politics" to "Very interested in politics."

The main predictor variable is "Q271" which asks “How important is being
\[RACIAL GROUP\] to your identity?”. The response options are a
five-point scale from "Not at all important" to "Extremely important".
Other potential predictors include survey questions on respondent's
self-perception of race and discrimination, and their own attitudes
towards other racial groups. In addition, there are multiple variables
that could be used as demographic controls such as age, gender,
education, income, and party identification. The variables not shown in
the output include responses to the rest of the 800 survey questions,
covering topics such as policy attitudes, civic engagement, and
political participation.

**Variable Type**: Ordinal

**Research question 2:** Controlling for other demographic variables
such as age and gender, how do experiences with racism influence the
importance respondents place on policy issues related to racial and
ethnic discrimination?

-   Outcome variable (include the name/description and type of
    variable): Stopping discrimination against racial/ethnic minorities
    (Q1R5) — a binary variable indicating whether respondents view it as
    an important issue.

**Variable Type:** Binary.

**Interaction Term Idea:** Include an interaction term between
race/ethnicity (S2RACER) and experiences with racism to see if the
influence of racism experiences differs across racial/ethnic groups.

**Load the data and provide a `glimpse()`**:

```{r data-1, warning=FALSE, echo=FALSE}
rm(list=ls())
load("39096-0001-Data.rda")
data <- da39096.0001
reordered_data <- data %>% relocate(Q29, Q271, WEIGHT, 
                                    AGE1_HID, AGE2_HID, .after = DATE)
reordered_data %>% slice_head(n = 50) %>% str()
```

\newpage

**Exploratory Plots:**
Dataset 1: Research Question 1
```{r plots-1.1, echo=FALSE, warning=FALSE}
data <- data %>% mutate(Q29_shortened = case_when(
  Q29 == "(1) Very interested in politics" ~ "Very interested",

  Q29 == "(2) Somewhat interested" ~ "Somewhat interested",

  Q29 == "(3) Not that interested in politics" ~ "Not that interested",

  Q29 == "(4) Not at all interested in politics" ~ 
    "Not at all interested", .default = "Other"))

wtd_tableQ29 <- data %>% count(Q29_shortened, wt=WEIGHT)
wtd_tableQ29 <- wtd_tableQ29 %>% mutate('pct' = (n/sum(n, na.rm=TRUE)))
ggplot(wtd_tableQ29, aes(x=Q29_shortened, y = pct)) +
  geom_bar(stat="identity", fill = "orange") +
  geom_text(aes(label=percent(pct)), vjust=1.6, color="white", size =3.5) +
  labs(title = "Reported Interest in Politics", 
       x = "Level of Interest", y = "Percent (Weighted)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

contingency_table <- table(data$Q29_shortened, data$Q271)
contingency_df <- as.data.frame(contingency_table)
contingency_df <- contingency_df %>% mutate(Share = (Freq/sum(Freq)))

ggplot(contingency_df, aes(x = Var1, y = Var2, fill = Share)) +
  geom_tile() +
  labs(title="Interest in Politics and Racial Identity", 
       x = "Interest in Politics", 
       y = "Importance of Racial Identity") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 0.4, vjust = 0.5)) +
  scale_fill_continuous(low = "white", high = "blue", labels = percent_format(scale = 100)) 
```

\newpage
Dataset 1: Research Question 2
```{r plots-1.2, echo=FALSE, warning=FALSE}

data$Q1R5_cleaned <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", data$Q1R5))
data$Q1R5_cleaned <- factor(data$Q1R5_cleaned, levels = c(0, 1), labels = c("Not Important", "Important"))

# Plot the distribution of Q1R5
ggplot(data, aes(x = Q1R5_cleaned)) +
  geom_bar(fill = "skyblue") +
  labs(title = "Importance of Stopping Racial Discrimination",
       x = "Importance Level",
       y = "Count") +
  theme_minimal()

data$MAIN_ETH_QUOTA_numeric <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", 
                                              data$MAIN_ETH_QUOTA))
data$Race_Ethnicity <- factor(data$MAIN_ETH_QUOTA_numeric,
                              levels = 1:7,
                              labels = c("White", "Latino", "Black", 
                                         "AAPI", "ANAI", "MENA", "NH"))

# Plot with short-term race labels on the x-axis and percent labels on the bars
ggplot(data, aes(x = Race_Ethnicity, fill = Q1R5_cleaned)) +
  geom_bar(position = "fill") +
  labs(title = "Importance of Stopping Discrimination by Race/Ethnicity",
       x = "Race/Ethnicity",
       y = "Proportion",
       fill = "Importance Level",
       caption = "Race/Ethnicity Abbreviations: ANAI = Native American, NH = Native Hawaiian, AAPI = Asian American/Pacific Islander, MENA = Middle Eastern/North African") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() 
```

\newpage

# Dataset 2

**Data source:** Global Health Observatory (GHO) data repository under
the World Health Organization (WHO), with additional economic data
sourced from the United Nations website.
https://www.kaggle.com/datasets/kumarajarshi/life-expectancy-who

**Brief description:** There are 2938 rows and 22 columns, representing
health and social/economic indicators for 193 countries over 15 years.
Each row corresponds to a specific country-year combination.

**Research question 1:** How does the number of years a person spends in school 
influence whether a country is classified as developed or developing?

-   Outcome variable (include the name/description and type of
    variable): "Status" --\> Binary variable that categorizes countries
    into "Developed" or "Developing."

**Research question 2:** How do factors such as life expectancy,
education, healthcare expenditure, and population affect GDP across
developing countries?

-   Outcome variable (include the name/description and type of
    variable): GDP: Numeric Variable

**Load the data and provide a `glimpse()`**:

```{r data-2, warning=FALSE, echo=FALSE}
data <- read.csv("Life Expectancy Data.csv")
glimpse(data)
```

\newpage

**Exploratory Plots:**
Dataset 2: Research Question 1
```{r plots-2.1, echo=FALSE, warning=FALSE}
# Bar chart showing the counts of each Status
ggplot(data, aes(x = Status, fill = Status)) +
  geom_bar(alpha = 0.8) +
  labs(
    title = "Distribution of Countries by Status",
    x = "Country Status",
    y = "Number of Countries",
    fill = "Country Status"
  ) +
  theme_minimal()

# Ridge plot for Schooling
ggplot(data, aes(x = Schooling, y = Status, fill = Status)) +
  geom_density_ridges(alpha = 0.7) +
  labs(
    title = "Distribution of years of schooling by Country Status",
    x = "Schooling",
    y = "Country Status"
  ) +
  theme_minimal()
```

\newpage
Dataset 2: Research Question 2
```{r plots-2.2, echo=FALSE, warning=FALSE}
developing_data <- subset(data, Status == "Developing")

# Histogram showing the distribution of GDP
ggplot(developing_data, aes(x = GDP)) +
  geom_histogram(binwidth = 100, 
                 fill = "skyblue", 
                 color = "black", 
                 alpha = 0.7) +
  labs(title = "Distribution of GDP",
       x = "GDP",
       y = "Frequency") + theme_minimal()

# 
ggplot(developing_data, aes(x = Schooling, y = GDP)) +
  geom_point(color = "blue", size = 2, alpha = 0.5) +
  labs(title = "Relationship Between years of education and GDP",
       x = "Years of education",
       y = "GDP") + theme_minimal()
```

\newpage

# Dataset 3 (optional)

**Data source:**
https://www.kaggle.com/datasets/prasoonkottarathil/polycystic-ovary-syndrome-pcos

**Brief description:** Each row in the dataset represents data of a
female. Specifically, it contains medical, physical, hormonal, and
lifestyle information collected for that woman as part of a study on
PCOS (Polycystic Ovary Syndrome) and related health metrics.

**Research question 1:** - What combination of physical characteristics
(e.g., BMI, waist-hip ratio), hormonal factors (e.g., AMH, FSH/LH
ratio), and symptoms (e.g., pimples, hair loss) significantly predict
PCOS status (Y/N)?

-   Outcome variable (include the name/description and type of
    variable): PCOS Status

Categorical (Binary: 1 = Has PCOS, 0 = Does not have PCOS) -
Description: Indicates whether the individual has PCOS or not.

**Research question 2:** How do physical factors and hormonal factors
levels influence Hb (hemoglobin) levels?

-   Outcome variable (include the name/description and type of
    variable): Hb (Hemoglobin) Levels

-   Type: Continuous (measured in g/dL)

-   Description:The concentration of hemoglobin in the blood, which can
    be an indicator of overall health, anemia, and other medical
    conditions.

**Load the data and provide a `glimpse()`**:

```{r data-3, warning=FALSE, echo=FALSE}
data <- read_excel("PCOS_data_without_infertility.xlsx", sheet = 2)
glimpse(data)
data <- data %>%
  rename(PCOS = `PCOS (Y/N)`, 
         Weight=`Weight (Kg)`, 
         Hb=`Hb(g/dl)`)
data$PCOS <- ifelse(data$PCOS == 1, "Yes", "No")
data$PCOS <- factor(data$PCOS, levels = c("No", "Yes"))
```

\newpage

**Exploratory Plots:**
Dataset 3: Research Question 1
```{r plots-3.1, echo=FALSE}
# Summarize data
pcos_summary <- data %>%
  count(PCOS)

# Create the pie chart
ggplot(pcos_summary, aes(x = "", y = n, fill = PCOS)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  labs(title = "PCOS Distribution") +
  theme_void() +
  scale_fill_manual(values = c("skyblue", "salmon"))

# Boxplot comparing Weight by PCOS status
ggplot(data, aes(x = PCOS, y = Weight, fill = PCOS)) +
  geom_boxplot() +
  labs(
    title = "Boxplot of Weight by PCOS Status",
    x = "PCOS Status (1 = Yes, 0 = No)",
    y = "Weight (kg)"
  ) +
  # Custom x-axis labels
  scale_x_discrete(labels = c("0" = "No", "1" = "Yes")) +  
  scale_fill_manual(values = c("#F8766D", "#00BFC4")) +    
  theme_minimal() +
  theme(legend.position = "none")
```

\newpage

Dataset 3: Research Question 2
```{r plots-3.2, echo=FALSE}
# Histogram of Hb levels
ggplot(data, aes(x = Hb)) +
  geom_histogram(binwidth = 0.5, 
                 fill = "#00BFC4", 
                 color = "black", 
                 alpha = 0.7) +
  labs(title = "Histogram of Hb (Hemoglobin) Levels",
       x = "Hemoglobin (g/dL)", 
       y = "Frequency") +
  theme_minimal()

# Line plot of BMI and Hb with regression line
ggplot(data, aes(x = `PRL(ng/mL)`, y = Hb)) +
  geom_point(color = "#00BFC4", size = 3) +  # Points for BMI vs Hb
  geom_smooth(method = "lm", color = "blue", se = FALSE) +
  labs(
    title = "Line Plot of PRL(ng/mL) vs Hb (Hemoglobin) Levels",
    x = "PRL(ng/mL)",
    y = "Hemoglobin (g/dL)") +
  theme_minimal()
```

\newpage

# Team Charter

#### When will you meet as a team to work on the project components? Will these meetings be held in person or virtually?
-   We will meet virtually (i.e through Zoom) every 7 days in the initial phase.
    However, near the deadline, we will aim to meet every 2 days and
    give updates on our individual tasks.

#### What is your group policy on missing team meetings (e.g., how much advance notice should be provided)?

-   Given that we are all juggling multiple projects/deadline, we can
    miss a meeting provided we give a 24 hour notice and follow-up with
    what we miss in the meetings and do the work that we are assigned

#### How will your team communicate (email, Slack, text messages)? What is your policy on appropriate response time (within a certain number of hours? Nights/weekends?)?

-   Our main form of communication would be WhatsApp.
-   We are not obligated to respond in the night, however if we are
    required to respond in the day, a half day buffer is acceptable